mruby Bytecode Loader Using Bluetooth in Multi-VM Environment

### 2016-04-06 EMSOFT2016/diff/0330_paper.tex
update

An approach of efficient software development is to use component-based techniques.
CBD (Component-Based Development) is a design technique for constituting reusable components \cite{par:Crnkovic}.
Complex and large scale software systems can be developed efficiently using component-based techniques.
That is because software componentization provides high-reusability and easy verification.
It also makes a system flexible in extensions and specification changes.
There are TECS \cite{par:TECS}, AUTOSAR \cite{url:AUTOSAR}, and SaveCCM \cite{par:SAVEapproach} as a typical component-based development for embedded systems

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/proposed.eps}
    \caption{System Model of the proposed framework}
    \label{fig:proposed}
\end{figure}

@@

mruby is suitable for embedded systems because of faster execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, therefore mruby programs can run on any operating system as long as VM is implemented.
A RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by a RiteVM.
The bytecodes can run on a Rite VM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby.eps}
    \caption{Mechanism of mruby/RiteVM}
    \label{fig:mruby}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_diagram.eps}
    \caption{Component Diagram}
    \label{fig:component}
\end{figure}

@@

\begin{description}
    \item[{\mybf Signature Description}]\mbox{}\\
        The {\myit signature} description defines an interface of a {\myit cell}.
        A {\myit signature} name is described following the keyword {\myit signature}.
        It also has the prefix ``s".
        In this way, a {\myit signature} is defined such as sMotor shown in Figure \ref{signature}.
        To make the definition of an interface clear, specifiers such as in and out are used in TECS.
        [in] and [out] represent input and output, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor{
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\caption{Signature Description}
\label{signature}
\end{figure}
    \item[{\mybf Celltype Description}]\mbox{}\\
        The {\myit celltype} description defines {\myit entry} ports, {\myit call} ports, attributes, and valuables of a {\myit celltype}.
        An example of a {\myit celltype} description is shown in Figure \ref{celltype}.
        A {\myit celltype} name following the keyword {\myit celltype} with the prefix ``t" and elements of a {\myit celltype} is described.
        To define {\myit entry} ports, a {\myit signature} such as sMotor, and an {\myit entry} port name such as eMotor follow the keyword {\myit entry}.
        In the same way, {\myit call} ports can be defined.
        Attributes and valuables follow the keyword {\myit attr} and {\myit var} respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller{
    call sMotor cMotor;
};
celltype tMotor{
    entry sMotor eMotor;
    attr{
        int32_t attribute;
    };
    var{
        int32_t variable;
    };
};
\end{lstlisting}
\caption{Celltype Description}
\label{celltype}
\end{figure}
    \item[{\mybf Build Description}]\mbox{}\\
        The build description is used to instantiate {\myit cell}s and connect {\myit cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\myit celltype} name such as tMotor, and a {\myit cell} name such as Motor follow the keyword {\myit cell}.
        To compose {\myit cell}s, a {\myit call} port, a {\myit signature}, a {\myit entry} port in order are described.
        In this example, a {\myit entry} port eMotor in a {\myit cell} Motor is connected to a {\myit call} port cMotor in a {\myit cell} Caller.
\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor{
};
cell tCaller Caller{
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\caption{Build Description}
\label{build}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mrubyontecs.eps}
    \caption{System Model of existing mruby on TECS}
    \label{fig:mrubyontecs}
\end{figure}

@@

Figure \ref{fig:mruby_TECS_bridge} shows an example of use of an mruby-TECS bridge for controlling a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby_TECS_bridge.eps}
    \caption{mruby-TECS Bridge}
    \label{fig:mruby_TECS_bridge}
\end{figure}

@@

\section{Design and Implementation}
\label{sec:Design and Implementation}
Figure \ref{fig:system_model} shows the detailed system model of the proposed framework.
Each mruby application bytecode transferred from the host is received by the loader in the RiteVM.
The RiteVM reads the own bytecode and executes it.
The mruby applications run at the same time because of synchronization processing.
The RiteVM scheduler switches tasks as multiple tasks can run in concurrent.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/system_model.eps}
    \caption{Detailed System Model of the proposed framework}
    \label{fig:system_model}
\end{figure}

@@

% \begin{figure}[t]
%     \centering
%     \includegraphics[width=8cm,clip]{figure/bluetooth_loader.eps}
%     \caption{Development Flow in mruby bytecode loader using Bluetooth}
%     \label{fig:bluetooth_loader}
% \end{figure}

\subsubsection{Component of RiteVM with mruby bytecode loader using Bluetooth}
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_bluetooth.eps}
    \caption{Component Diagram of mruby bytecode loader using Bluetooth}
    \label{fig:component_bluetooth}
\end{figure}
The proposed framework provides a RiteVM with mruby bytecode loader using Bluetooth as TECS component.
The component is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages a RiteVM configuration such as automatically generation the mruby library bytecode.
This generated bytecode is prepared beforehand in the storage/ROM device, and different from a bytecode transferred with Bluetooth.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/control_flow.eps}
    \caption{Process Flow of mruby bytecode loader using Bluetooth}
    \label{fig:control_flow}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.cdl */
celltype tRiteVMBluetooth{
    entry sTaskBody eMrubyBody;
    [optional] call sEventflag cEventflag[];
    [optional] call sSemaphore cSemaphore;
    attr{
      [omit]char_t *mrubyLib;
      char_t *irepLib = 
                    C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = 
                    C_EXP( BUFFER_SIZE );
      FLGPTN setptn;
    };
    var{
        mrb_state *mrb;
        mrbc_context *context;
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\caption{Celltype Description for RiteVM with mruby bytecode loader using Bluetooth}
\label{celltype_mrubybluetooth}
\end{figure}

@@

First, the mruby bytecode loader receives the mruby application bytecode transferred from the host (Figure \ref{fig:control_flow}:A, Figure \ref{maincode_mrubybluetooth}:6-7).
The bytecode is stored in the component variable such as VAR\_irepApp shown in Figure \ref{celltype_mrubybluetooth}.
This process is exclusively carried out by the semaphore.

Second, pointers of {\myit mrb\_state} and {\myit mrbc\_context}, and mruby-TECS bridges are initialized (Figure \ref{fig:control_flow}:B, Figure \ref{maincode_mrubybluetooth}:9-14).
{\myit mrb\_state} is a set of states and global variables used in mruby.
Synchronization of multiple tasks is performed in this processing phase.
The RiteVM that ended up here waits for the other RiteVM to finish loading and initialization.

Third, the RiteVM reads the bytecode of mruby libraries (Figure \ref{fig:control_flow}:C, Figure \ref{maincode_mrubybluetooth}:17-19).
mruby libraries are a set of Ruby classes such as motor class and sensor class.
For example, motor class defines methods to rotate and stop a motor.
The tRiteVMBluetooth {\myit cell} has attributes as shown in Figure \ref{celltype_mrubybluetooth}.
The {\myit mrubyLib} indicates the program files of mruby libraries.
{\myit [omit]} is only used for the TECS generator, thus the attribute, {\myit mrubyLib}, does not consume memory.
The {\myit irepLib} is the pointer of the array stored the bytecode of mruby libraries.
In short, the bytecode of mruby libraries is stored as an attribute of the component when compiling for the first time.

Fourth, the RiteVM reads the bytecode of the mruby application transferred with Bluetooth (Figure \ref{fig:control_flow}:D, Figure \ref{maincode_mrubybluetooth}:20-22).
The mruby application bytecode is stored in an array of {\myit irepApp}.
The array is different from that of holding the mruby library bytecode.
Two bytecodes are read separately in the RiteVM.

@@

The process shown in Figure \ref{maincode_mrubybluetooth}:23-28 is carried out when an exception occurs.
When all mruby applications finish, mrb\_state and mrbc\_context are freed (Figure \ref{maincode_mrubybluetooth}:31-34).
The proposed framework supports continuous loading, thus this process loops.
After freeing variables, RiteVM will become the state to wait for the next mruby application bytecode.
\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.c */
void
eMrubyBody_main( CELLIDX idx )
{
  /* Omit: start of exclusive process by semaphore */
  /* Receive the bytecode via Bluetooth */
  bluetooth_loader( VAR_irepApp );
  /* Omit: end of exclusive process by semaphore */
  /* New interpreter instance */
  VAR_mrb = mrb_open();
  /* Omit: error check for mrb_state */
  /* New mruby context */
  VAR_context = mrbc_context_new( VAR_mrb );
  /* Omit: initialization of mruby-TECS bridge */
  /* Omit: synchronization of
                starting mruby application */
  /* Load mruby library bytecode */
  mrb_load_irep_cxt( VAR_mrb,
                        ATTR_irepLib, VAR_context );
  /* Load mruby application bytecode and run */
  mrb_load_irep_cxt( VAR_mrb,
                        VAR_irepApp, VAR_context );
  if ( mrb->exc ) {
    /* Failure to execute */
    mrb_p( VAR_mrb, 
                mrb_obj_value( VAR_mrb->exc ) );
    exit( 0 );
  }
  /* Omit: synchronization of
                finishing mruby application */
  /* Free mruby context */
  mrbc_context_free( VAR_mrb, VAR_context );
  /* Free interpreter instance */
  mrb_close( VAR_mrb );
}

\end{lstlisting}
\caption{Main code for RiteVM with mruby bytecode loader using Bluetooth}
\label{maincode_mrubybluetooth}
\end{figure}
\subsection{RiteVM Scheduler}
\label{sec:RiteVM Scheduler}
This section describes implementation of RiteVM scheduler in the proposed framework.
mruby on TECS has supported multitasking.
However, multitask processing in mruby on TECS requires the knowledge of the RTOS (TOPPERS/HRP2) for developers.

@@

\subsubsection{Design of RiteVM Scheduler}
A RiteVM scheduler is a periodic handler, and calls {\myit rotateReadyQueue}, a service call of $\mu$ITRON to switch tasks with the same priority.
In other words, the RiteVM scheduler calls {\myit rotateReadyQueue} cyclically.
The design of the RiteVM scheduler is shown in Figure \ref{fig:rotateReadyQueue}. 
{\myit rotateReadyQueue} is described as follows.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/rotateReadyQueue.eps}
    \caption{The design of RiteVM scheduler}
    \label{fig:rotateReadyQueue}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/cyclic_handler.eps}
    \caption{Component Diagram of RiteVM Scheduler}
    \label{fig:cyclic_handler}
\end{figure}
\begin{figure}[t]
    \centering
    \begin{lstlisting}
/* tRiteVMScheduler.cdl */
celltype tCyclicHandler {
    [inline] entry sCyclic eCyclic;
    call siHandlerBody  ciBody;
    attr {
        [omit] ATR    attribute   = C_EXP("TA_NULL");
    	[omit] RELTIM cyclicTime;
    	[omit] RELTIM cyclicPhase;
    };
};
celltype tRiteVMSchedulerMain {
    require tKernel.eiKernel;
    entry   siHandlerBody eiBody;
    attr {
        PRI priority;
    };
};

composite tRiteVMScheduler {
    attr {
        ATR    attribute      = C_EXP("TA_NULL");
        RELTIM cyclicTime  = 1;
        RELTIM cyclicPhase = 1;
        PRI    priority;
    };
    cell tRiteVMSchedulerMain RiteVMSchedulerMain {
        priority = composite.priority;
    };
    cell tCyclicHandler CyclicHandler {
        ciBody      = RiteVMSchedulerMain.eiBody;
        attribute   = composite.attribute;
        cyclicTime  = composite.cyclicTime;
        cyclicPhase = composite.cyclicPhase;
    };
};
    \end{lstlisting}
    \caption{Celltype Description of RiteVM Scheduler}
    \label{celltype_cyclic_handler}
\end{figure}
\begin{figure}[t]
    \centering
    \begin{lstlisting}
cell tRiteVMScheduler RiteVMScheduler {
    attribute   = C_EXP("TA_STA");
    cyclicTime  = 1;
    cyclicPhase = 1;
    priority    =
        C_EXP("RITEVM_PRIORITY");
};
\end{lstlisting}
    \caption{Build Description of RiteV Scheduler}
    \label{build_cyclic_handler}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/Eventflag.eps}
    \caption{The design for Eventflag using TECS {\scriptsize *only differences}}
    \label{fig:Eventflag}
\end{figure}

@@

\section{Experimental Evaluation}
\label{sec:Evaluation}
This section mentions experimental results and their consideration.
To analyze the advantages of the proposed framework, the evaluations are performed as follows.
\begin{itemize}
%    \item Execution time of the platform
    \item Size and time for transferred mruby bytecodes
    \item Execution time with singletasking, co-routine, and multitasking
    \item Overhead for periodic time
%    \item Synchronization of mruby applications
    \item Code size utilizing component-based development 
\end{itemize}

@@

\subsection{Improving the software development efficiency by mruby bytecode loader}
The size, load process time, and compilation time for mruby application and mruby application including library is shown in Table \ref{tab:size_and_time}.
The overhead of load processing is 50.933 msec, which it takes to load a zero bytes bytecode.
Similarly, the overhead of compilation is 46.9 msec, which it takes to compile a zero bytes program.
The mruby application bytecode is smaller and faster than that of including mruby libraries in all terms.
The difference becomes larger as the number of RiteVMs increases because it takes the 50 msec overhead per one RiteVM. 
%In the proposed framework, developers send only mruby application and prepare mruby library in advance.
%Therefore, the design can send the bytecode faster.
In addition, the design can save the time to rewrite a storage/ROM device, and to restart an OS.
These advantages lead improvement of the software development efficiency.

\begin{table}[t]
    \centering
    \caption{Comparison of the size and load process time between an mruby application including mruby libraries and not}
    \scriptsize
    {\tabcolsep=0.1cm
    \begin{tabular}{c||c|c|c}
                            & App\&Lib     & App        &   App\&Lib/App  \\ \hline
          Bytecode Size     & 14,044 bytes & 199 bytes  &   $\times$70.6          \\ %\hline
          Loading Time      & 305.081 msec & 7.774 msec &   $\times$39.2          \\
          Compilation Time  & 8.7 msec     & 0.3 msec   &   $\times$29.0          \\
    \end{tabular}
    }
    \label{tab:size_and_time}
\end{table}

>>>>

An approach of efficient software development is to use component-based techniques.
CBD (Component-Based Development) is a design technique for constituting reusable components \cite{par:Crnkovic}.
Complex and large scale software systems can be developed efficiently using component-based techniques.
That is because software componentization provides high-reusability.
Verification of component-based systems has been researched \cite{par:Blaming} \cite{par:Verification}.
It also makes a system flexible in extensions and specification changes.
There are TECS \cite{par:TECS}, AUTOSAR \cite{url:AUTOSAR}, and SaveCCM \cite{par:SAVEapproach} as a typical component-based development for embedded systems

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/proposed.eps}
    \vspace{1mm}
    \caption{System Model of the proposed framework}
    \vspace{1mm}
    \label{fig:proposed}
\end{figure}

@@

mruby is suitable for embedded systems because of execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, therefore mruby programs can run on any operating system as long as VM is implemented.
A RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by a RiteVM.
The bytecodes can run on a Rite VM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby.eps}
    \vspace{1mm}
    \caption{Mechanism of mruby/RiteVM}
    \vspace{1mm}
    \label{fig:mruby}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_diagram.eps}
    \vspace{1mm}
    \caption{Component Diagram}
    \vspace{1mm}
    \label{fig:component}
\end{figure}

@@

\begin{description}
    \item[{\mybf Signature Description}]\mbox{}\\
        The {\myit signature} description defines an interface of a {\myit cell}.
        A {\myit signature} name is described following the keyword {\myit signature}.
        It also has the prefix ``s".
        In this way, a {\myit signature} is defined such as sMotor shown in Figure \ref{signature}.
        To make the definition of an interface clear, specifiers such as in and out are used in TECS.
        [in] and [out] represent input and output, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor{
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\vspace{1mm}
\caption{Signature Description}
\vspace{1mm}
\label{signature}
\end{figure}
    \item[{\mybf Celltype Description}]\mbox{}\\
        The {\myit celltype} description defines {\myit entry} ports, {\myit call} ports, attributes, and valuables of a {\myit celltype}.
        An example of a {\myit celltype} description is shown in Figure \ref{celltype}.
        A {\myit celltype} name following the keyword {\myit celltype} with the prefix ``t" and elements of a {\myit celltype} is described.
        To define {\myit entry} ports, a {\myit signature} such as sMotor, and an {\myit entry} port name such as eMotor follow the keyword {\myit entry}.
        In the same way, {\myit call} ports can be defined.
        Attributes and valuables follow the keyword {\myit attr} and {\myit var} respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller{
    call sMotor cMotor;
};
celltype tMotor{
    entry sMotor eMotor;
    attr{
        int32_t port;
    };
    var{
        int32_t currentSpeed = 0;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype Description}
\vspace{1mm}
\label{celltype}
\end{figure}
    \item[{\mybf Build Description}]\mbox{}\\
        The build description is used to instantiate {\myit cell}s and connect {\myit cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\myit celltype} name such as tMotor, and a {\myit cell} name such as Motor follow the keyword {\myit cell}.
        To compose {\myit cell}s, a {\myit call} port, a {\myit signature}, a {\myit entry} port in order are described.
        In this example, a {\myit entry} port eMotor in a {\myit cell} Motor is connected to a {\myit call} port cMotor in a {\myit cell} Caller.
\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor{
};
cell tCaller Caller{
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\vspace{1mm}
\caption{Build Description}
\vspace{1mm}
\label{build}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mrubyontecs.eps}
    \vspace{1mm}
\caption{System Model of existing mruby on TECS}
    \vspace{1mm}
\label{fig:mrubyontecs}
\end{figure}

@@

Figure \ref{fig:mruby_TECS_bridge} shows an example of use of an mruby-TECS bridge for controlling a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby_TECS_bridge.eps}
    \vspace{1mm}
\caption{mruby-TECS Bridge}
    \vspace{1mm}
\label{fig:mruby_TECS_bridge}
\end{figure}

@@

\section{Design and Implementation}
\label{sec:Design and Implementation}
Figure \ref{fig:system_model} shows the detailed system model of the proposed framework.
Each mruby application bytecode transferred from the host to the target deviceis received by the loader in the RiteVM.
The RiteVM reads the transferred bytecode and executes it with libraries.
The mruby applications run at the same time because of synchronization processing.
The RiteVM scheduler switches RiteVM tasks as multiple tasks can run in concurrent.
This section explains these proposed functionalities.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/system_model.eps}
    \vspace{1mm}
\caption{Detailed System Model of the proposed framework}
    \vspace{1mm}
\label{fig:system_model}
\end{figure}

@@

% \begin{figure}[t]
%     \centering
%     \includegraphics[width=8cm,clip]{figure/bluetooth_loader.eps}
%     \vspace{1mm}
% \caption{Development Flow in mruby bytecode loader using Bluetooth}
%     \vspace{1mm}
% \label{fig:bluetooth_loader}
% \end{figure}

\subsubsection{Component of RiteVM with mruby bytecode loader using Bluetooth}
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_bluetooth.eps}
    \vspace{1mm}
\caption{Component Diagram of mruby bytecode loader using Bluetooth}
    \vspace{1mm}
\label{fig:component_bluetooth}
\end{figure}
The proposed framework provides a RiteVM with mruby bytecode loader using Bluetooth as TECS component.
The component is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages a RiteVM configuration such as automatically generation the mruby library bytecode.
This generated bytecode is prepared beforehand in the storage/ROM device, and different from a bytecode transferred with Bluetooth.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/control_flow.eps}
    \vspace{1mm}
\caption{Process Flow of mruby bytecode loader using Bluetooth}
    \vspace{1mm}
\label{fig:control_flow}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.cdl */
celltype tRiteVMBluetooth{
    entry sTaskBody eMrubyBody;
    [optional] call sEventflag cEventflag[];
    [optional] call sSemaphore cSemaphore;
    attr{
      [omit]char_t *mrubyLib;
      char_t *irepLib = 
                    C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = 
                    C_EXP( BUFFER_SIZE );
      FLGPTN setptn;
    };
    var{
        mrb_state *mrb;
        mrbc_context *context;
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\vspace{1mm}
\caption{Celltype Description for RiteVM with mruby bytecode loader using Bluetooth}
\vspace{1mm}
\label{celltype_mrubybluetooth}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.c */
void
eMrubyBody_main( CELLIDX idx )
{
  /* Omit: start of exclusive process by semaphore */
  /* Receive the bytecode via Bluetooth */
  bluetooth_loader( VAR_irepApp );
  /* Omit: end of exclusive process by semaphore */
  /* New interpreter instance */
  VAR_mrb = mrb_open();
  /* Omit: error check for mrb_state */
  /* New mruby context */
  VAR_context = mrbc_context_new( VAR_mrb );
  /* Omit: initialization of mruby-TECS bridge */
  /* Omit: synchronization of
                starting mruby application */
  /* Load mruby library bytecode */
  mrb_load_irep_cxt( VAR_mrb,
                        ATTR_irepLib, VAR_context );
  /* Load mruby application bytecode and run */
  mrb_load_irep_cxt( VAR_mrb,
                        VAR_irepApp, VAR_context );
  if ( mrb->exc ) {
    /* Failure to execute */
    mrb_p( VAR_mrb, 
                mrb_obj_value( VAR_mrb->exc ) );
    exit( 0 );
  }
  /* Omit: synchronization of
                finishing mruby application */
  /* Free mruby context */
  mrbc_context_free( VAR_mrb, VAR_context );
  /* Free interpreter instance */
  mrb_close( VAR_mrb );
}

\end{lstlisting}
\vspace{1mm}
\caption{Main code for RiteVM with mruby bytecode loader using Bluetooth}
\vspace{1mm}
\label{maincode_mrubybluetooth}
\end{figure}

@@

First, the mruby bytecode loader receives the mruby application bytecode transferred from the host (Figure \ref{fig:control_flow}:A, Lines 6-7 in Figure \ref{maincode_mrubybluetooth}).
The bytecode is stored in the component variable such as VAR\_irepApp shown in Figure \ref{celltype_mrubybluetooth}.
This process is exclusively carried out by the semaphore.

Second, pointers of {\myit mrb\_state} and {\myit mrbc\_context}, and mruby-TECS bridges are initialized (Figure \ref{fig:control_flow}:B, Lines 9-14 in Figure \ref{maincode_mrubybluetooth}).
{\myit VAR\_mrb} and {\myit VAR\_context} shows the variable of the {\myit cell}.
{\myit mrb\_state} is a set of states and global variables used in mruby.
Synchronization of multiple tasks is performed in this processing phase.
The RiteVM that ended up here waits for the other RiteVM to finish loading and initialization.

Third, the RiteVM reads the bytecode of mruby libraries (Figure \ref{fig:control_flow}:C, Lines 17-19 in Figure \ref{maincode_mrubybluetooth}).
mruby libraries are a set of Ruby classes such as motor class and sensor class.
For example, motor class defines methods to rotate and stop a motor.
The tRiteVMBluetooth {\myit cell} has attributes as shown in Figure \ref{celltype_mrubybluetooth}.
{\myit ATTR} shows the attribute, which is a fixed value and not rewritten unlike {\myit VAR}.
The {\myit mrubyLib} indicates the program files of mruby libraries.
{\myit mrubyLib} is the attribute because mruby libraries are not modified in the proposed development process.
{\myit [omit]} is only used for the TECS generator, thus the attribute, {\myit mrubyLib}, does not consume memory.
The {\myit irepLib} is the pointer of the array stored the bytecode of mruby libraries.
In short, the bytecode of mruby libraries is stored as an attribute of the component when compiling for the first time.

Fourth, the RiteVM reads the bytecode of the mruby application transferred with Bluetooth (Figure \ref{fig:control_flow}:D, Lines 20-22 in Figure \ref{maincode_mrubybluetooth}).
The mruby application bytecode is stored in an array of {\myit irepApp}.
The array is different from that of holding the mruby library bytecode.
Two bytecodes are read separately in the RiteVM.

@@

The process shown in Figure \ref{maincode_mrubybluetooth}:23-28 is carried out when an exception occurs.
When all mruby applications finish, mrb\_state and mrbc\_context are freed (Figure \ref{maincode_mrubybluetooth}:31-34).
The proposed framework supports continuous loading, thus this process loops.
After freeing variables, RiteVM will become the state to wait for the next mruby application bytecode.
\subsection{RiteVM Scheduler}
\label{sec:RiteVM Scheduler}
This section describes implementation of RiteVM scheduler in the proposed framework.
mruby on TECS has supported multitasking.
However, multitask processing in mruby on TECS requires the knowledge of the RTOS (TOPPERS/HRP2) for developers.

@@

The RiteVM scheduler is a fair scheduler and implemented as a TECS component in the proposed framework.
Therefore, when developers create software in priority-based scheduling, the RiteVM scheduler can be easily removed.
\subsubsection{Design of RiteVM Scheduler}
A RiteVM scheduler is a periodic handler, and calls {\myit rotateReadyQueue}, a service call of $\mu$ITRON to switch tasks with the same priority.
In other words, the RiteVM scheduler calls {\myit rotateReadyQueue} cyclically.
The design of the RiteVM scheduler is shown in Figure \ref{fig:rotateReadyQueue}. 
{\myit rotateReadyQueue} is described as follows.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/rotateReadyQueue.eps}
    \vspace{1mm}
\caption{The design of RiteVM scheduler}
    \vspace{1mm}
\label{fig:rotateReadyQueue}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/cyclic_handler.eps}
    \vspace{1mm}
\caption{Component Diagram of RiteVM Scheduler}
    \vspace{1mm}
\label{fig:cyclic_handler}
\end{figure}
\begin{figure}[t]
    \centering
    \begin{lstlisting}
/* tRiteVMScheduler.cdl */
celltype tCyclicHandler {
    [inline] entry sCyclic eCyclic;
    call siHandlerBody  ciBody;
    attr {
        [omit] ATR    attribute   = C_EXP("TA_NULL");
    	[omit] RELTIM cyclicTime;
    	[omit] RELTIM cyclicPhase;
    };
};
celltype tRiteVMSchedulerMain {
    require tKernel.eiKernel;
    entry   siHandlerBody eiBody;
    attr {
        PRI priority;
    };
};

composite tRiteVMScheduler {
    attr {
        ATR    attribute      = C_EXP("TA_NULL");
        RELTIM cyclicTime  = 1;
        RELTIM cyclicPhase = 1;
        PRI    priority;
    };
    cell tRiteVMSchedulerMain RiteVMSchedulerMain {
        priority = composite.priority;
    };
    cell tCyclicHandler CyclicHandler {
        ciBody      = RiteVMSchedulerMain.eiBody;
        attribute   = composite.attribute;
        cyclicTime  = composite.cyclicTime;
        cyclicPhase = composite.cyclicPhase;
    };
};
    \end{lstlisting}
    \vspace{1mm}
\caption{Celltype Description of RiteVM Scheduler}
    \vspace{1mm}
\label{celltype_cyclic_handler}
\end{figure}
\begin{figure}[t]
    \centering
    \begin{lstlisting}
cell tRiteVMScheduler RiteVMScheduler {
    attribute   = C_EXP("TA_STA");
    cyclicTime  = 1;
    cyclicPhase = 1;
    priority    =
        C_EXP("RITEVM_PRIORITY");
};
\end{lstlisting}
    \vspace{1mm}
\caption{Build Description of RiteV Scheduler}
    \vspace{1mm}
\label{build_cyclic_handler}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/Eventflag.eps}
    \vspace{1mm}
\caption{The design for Eventflag using TECS (only differences)}
    \vspace{1mm}
\label{fig:Eventflag}
\end{figure}

@@

\section{Experimental Evaluation}
\label{sec:Evaluation}
This section mentions experimental results and their consideration.
To analyze the advantages of the proposed framework, the evaluations are performed as follows.
\begin{itemize}
%    \item Execution time of the platform
    \item Size and time for transferred mruby bytecodes
    \item Execution time with singletasking, co-routine, and proposed multitasking
    \item Overhead for periodic time
    \item Synchronization of multiple RiteVM tasks
    \item Code size utilizing component-based development 
\end{itemize}

@@

\subsection{Improving the software development efficiency by mruby bytecode loader}
The mruby bytecode loader using Bluetooth can save time to develop software.
In the proposed framework, developers do not need to rewrite a storage/ROM device because the bytecode should be just transferred.
For example, when mruby programs are modified, developers must pull out an SD card from the target device, connect the host PC, compile/link the platform, and insert the SD card in the target device, in the experimental environment such as LEGO MINDSTORMS EV3.
In addtion, it can save the time to restart a target OS.
\subsubsection{Size and time for transferred mruby bytecodes}
In the framework, to further improve the software development efficiency, developers transfers only the mruby application bytecode and mruby libraries are incorporated in the platform.
The size, load process time, and compilation time for mruby application and mruby application including library is shown in Table \ref{tab:size_and_time}.
The overhead of load processing is 50.933 msec, which it takes to load a zero bytes bytecode.
Similarly, the overhead of compilation is 46.9 msec, which it takes to compile a zero bytes program.
The mruby application bytecode is smaller and faster than that of including mruby libraries in all terms.
The difference becomes larger as the number of RiteVMs increases because it takes the 50 msec overhead per one RiteVM. 
%In the proposed framework, developers send only mruby application and prepare mruby library in advance.
%Therefore, the design can send the bytecode faster.
%In addition, the design can save the time to rewrite a storage/ROM device, and to restart an OS.
These advantages lead improvement of the software development efficiency.

\begin{table}[t]
    \centering
    \vspace{1mm}
\caption{Comparison of the size and load process time between an mruby application including mruby libraries and not}
    \vspace{1mm}
    %\scriptsize
    {\tabcolsep=0.1cm
    \begin{tabular}{c||c|c|c}
                            & App\&Lib     & App        &   App\&Lib/App  \\ \hline
          Bytecode Size     & 14,044 bytes & 199 bytes  &   $\times$70.6          \\ %\hline
          Loading Time      & 305.081 msec & 7.774 msec &   $\times$39.2          \\
          Compile Time  & 8.7 msec     & 0.3 msec   &   $\times$29.0          \\
    \end{tabular}
    }
    \label{tab:size_and_time}
\end{table}
